---
title: "Outliers-health metadata"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, warning=FALSE, echo=FALSE, message=FALSE}
# loads the libraries
library(tidyverse)
library(knitr)
library(dplyr)
library(reshape2)

# import tables of the questionnaires and the fields informations
field_data <- read_csv("../Health_data/fields_extracted.csv")
TMonths_data <- read_csv("../Health_data/Health_3months_18.03.21.csv")
Reg_data <- read_csv("../Health_data/Health_Reg_18.03.21.csv")
```

for all int data, I want to plot the x--range of data, y--frequency of data

```{r}
# select the int fields from the field_data description
fields_int_3M <- field_data %>% filter(data_type=="int", file_name=="3Months") %>% select(field) %>% pull
fields_int_re <- field_data %>% filter(data_type=="int", file_name=="Regular") %>% select(field) %>% pull

TMonths_int <- TMonths_data %>% select(Family_ID, any_of(fields_int_3M))
Reg_int <- Reg_data %>% select(Family_ID, any_of(fields_int_re))
```
Here I used a for loop to plot each int data
```{r}

for (i in 1:13) {
  n <- TMonths_int %>% select(fields_int_3M[i]) %>% t()
  hist(n, freq = T, xlab = fields_int_3M[i], main = fields_int_3M[i])
}
```
14th inf_NbColicHospitalization have no valid data (1 for 0 and rest are NA)

here is the rest
```{r}
for (i in 15:29) {
  n <- TMonths_int %>% select(fields_int_3M[i]) %>% t()
  hist(n, freq = T, xlab = fields_int_3M[i], main = fields_int_3M[i])
}
```
for inf_SeriousIllness, there seems some misunderstand. Some parents consider the general health is 100, but the child
did have some disease which might be cured later. And they consider the serious illness scale is 0(which means they
believe the child had serous illness.)


here is the reg_data

```{r}

for (i in 1:length(fields_int_re)) {
  n <- Reg_int %>% select(fields_int_re[i]) %>% t()
  hist(n, freq = T, xlab = fields_int_re[i], main = fields_int_re[i])
}
```

