---
title: "Metaphlan_Health_3"
output: html_notebook
---
This scrip merged health data with metaphlan3 data

#Create Physeq
##TAX and OTU
```{r warning=FALSE, include=FALSE}
# create Phylloseq object
library(ape)
library(dplyr)
library(phyloseq)
library(tidyverse)
library(microbiome)
library(vegan)
library(ggplot2)

# import tables and format for PhylloSeq
metaphlan <- read_tsv("../Metaphlan3/merged_abundance_table.txt") 
# select the species level information
metaphlan_s <- metaphlan %>% separate(clade_name,  #separate clade_name into 7 columns
                                    into=c("kingdom","Phylum","class","order","family","genera","species"), 
                                    sep="\\|") %>%
                            separate(NCBI_tax_id, #separete NCBI_tax_id into 7 columns
                                    into=c("Tax_kingdom","Tax_Phylum","Tax_class","Tax_order","Tax_family","Tax_genera","TaxID"), 
                                    sep="\\|") %>%
                            mutate(kingdom=str_replace(kingdom, "k__", "")) %>%
                            mutate(Phylum=ifelse(kingdom=="Unknown", "Unknown", str_replace(Phylum, "p__", ""))) %>%
                            mutate(class=ifelse(kingdom=="Unknown", "Unknown", str_replace(class, "c__", ""))) %>%
                            mutate(order=ifelse(kingdom=="Unknown", "Unknown", str_replace(order, "o__", ""))) %>%
                            mutate(family=ifelse(kingdom=="Unknown", "Unknown", str_replace(family, "f__", ""))) %>%
                            mutate(genera=ifelse(kingdom=="Unknown", "Unknown", str_replace(genera, "g__", ""))) %>%
                            mutate(species=ifelse(kingdom=="Unknown", "Unknown", str_replace(species, "s__", ""))) %>%
                            mutate(TaxID=ifelse(kingdom=="Unknown", "-1", TaxID)) %>%
                            #columns from clade_name: keep only the tax-name and trim the "k__" prefix
                            filter(!is.na(species))
                            #filter out the rows that have no species information.

# remove samples with less than 10 taxonomic assignments
nonzero <- function(x) sum(x != 0)
meta_exclude <- metaphlan_s %>% summarise(across(starts_with("Hel"), nonzero)) %>% 
  pivot_longer(everything(),names_to="samples", values_to="nb_tax") %>%
  filter(nb_tax<10) %>% select(samples) %>%
  pull()

metaphlan_s <- metaphlan_s %>% select(- all_of(meta_exclude))

## create TaxTable
taxMat <-  metaphlan_s %>% select(TaxID, kingdom, Phylum, class, order, family, genera, species) 
taxMat <- taxMat %>% tibble::column_to_rownames("species")
TAX = tax_table(as.matrix(taxMat))

## create countTable
countMat <- metaphlan_s %>% select(-kingdom, -Phylum, -class, -order, -family, -genera, -TaxID, 
                                   -Tax_kingdom, -Tax_Phylum, -Tax_class, -Tax_order, -Tax_family, -Tax_genera)
countMat <- countMat %>% tibble::column_to_rownames("species")
OTU = otu_table(as.matrix(countMat), taxa_are_rows = TRUE)
```

Now we have TAX OTU but no SAMPLE.
##Create SAMPLE
```{r include=FALSE}
## sample description file
sampleMat <- read_csv("../Metaphlan3/sample_table_12.04.21.csv")

health_3m_meta <- read.csv("../Health_data/Health_3months_nonOutliers_ABscaled_17.04.21.csv")
health_Reg_meta <- read.csv("../Health_data/Health_Reg_nonOutliers_ABscaled_17.04.21.csv")

fieldsinfo <- read.csv("../Health_data/fields_usable_21.04.15.csv")

#Here I want to add all fields except "text" type fields from health_data into SAMPLE
#For late analysis, use info from 'fieldsinfo' to cluster the fields groups

fields_text <- fieldsinfo %>% filter(data_type=="text") %>% select(field) %>% pull
health_3m <- health_3m_meta %>% select(!any_of(fields_text))
health_Reg <- health_Reg_meta %>% select(!any_of(fields_text))

sample_health_3m <- left_join(sampleMat, health_3m, by = c("Family_ID", "sample_type"))
sample_health_Reg <- left_join(sampleMat, health_Reg, by = c("Family_ID", "sample_type"))
```

###Merge 3M data

The number of obs. of sample_health_3m_Mat should be the same as sampleMat(374)
```{r include=FALSE}
#Valid_from for B8 and B9
#B8: 52(keep) 65
#B9: 78(keep) 91
sample_health_3m_Mat <- sample_health_3m %>% filter(!(Valid_From %in% c(65,91)))
```

###Merge Reg data
To delete the rows with the same familyID and time code,
we need to create a list contain all suspect rows by Valid_from/to
and a list without any suspect rows called cleaned list.

Some not wanted rows might remain in cleaned list. And some are misfiltered for having the same Valid_from/to.

Then with anti_join we are able to gain a new list having misfiltered rows.

Finally, we can merge the misfiltered rows with the cleaned list by full_join.

The number of obs. of sample_health_Reg_Mat should be the same as sampleMat(374)

Note: in Reg_data, there are more than 2 B1 for one family.
And we need to treat mis_sample as the samole_health_Reg with the valide_from of secondary priority.
Thus the code here is not perfect but works for this job.

```{r include=FALSE}
#Valid_from for each time code
#Priority B1 B2 B3 B4 B5 B6 B7 B8 B9 F M
#1st      2 5 8 11 22 30 43 48 NA NA NA
#2nd      1 4 7 10 20 26 39 NA NA NA NA
#3rd      0 3 6 9 16 24 35 NA NA NA NA

clean_sample <- sample_health_Reg %>% filter(Valid_From %in% c(2,5,8,11,22,30,43,48))
suspect_sample <- sample_health_Reg %>% filter(!(Valid_From %in% c(2,5,8,11,22,30,43,48)))
mis_sample <- anti_join(suspect_sample,clean_sample, by = c("Family_ID", "sample_type"))

#filter mis_sample with 2nd priority valid_from
mis_clean <- mis_sample %>% filter(Valid_From %in% c(1, 4, 7, 10, 20, 26, 39))
mis_sus <- mis_sample %>% filter(!Valid_From %in% c(1, 4, 7, 10, 20, 26, 39))
mis_mis <- anti_join(mis_sus, mis_clean, by = c("Family_ID", "sample_type") )
mis_filtered <- full_join(mis_clean, mis_mis)

sample_health_Reg_Mat <- full_join(clean_sample, mis_filtered)
#the number of obs. of sample_health_Reg_Mat is 376, which means there are two extra rows.
#The extra rows can be found manually
#Family *** have three B5 valid from 18,13,12
#here use valid_from == 18
sample_health_Reg_Mat <- sample_health_Reg_Mat %>% filter(!(Family_ID=="***" & Valid_From %in% c(13,12)))

```

###Merge 3m and Reg
Now merge 3m_Mat and Reg_Mat to creat SAMPLE
```{r}
#Merge both 3m_Mat and Reg_Mat
sample_health <- left_join(sample_health_3m_Mat, sample_health_Reg_Mat, by = c("Family_ID", "sample_type","sample_ID","file_name","inf_DeliveryMode", "Run_ID"))

#creat pre-SAMPE table
sample_health <- sample_health %>% tibble::column_to_rownames("sample_ID")
```

###Filter out NA fields
There are some fields contain too many NA values. Here plot bool, AB and categorical fields to filter out these fields
```{r}
#bool
fields_bool <- fieldsinfo %>% filter(data_type=="bool") %>% select(field) %>% pull
sample_health_bool <- sample_health %>% select(Family_ID, any_of(fields_bool))

# get the counts for true, false and NA answers for each questions
sample_health_bool_long <- sample_health_bool %>% pivot_longer(!Family_ID, names_to="question", values_to="answers")
sample_health_bool_counts <- sample_health_bool_long %>% group_by(question, answers) %>% tally()

# plot the distribution
plotbools <- sample_health_bool_counts %>% ggplot(aes(x=question, y=n, fill=answers)) +
  geom_bar(stat="identity")+
  coord_flip() +
  theme(legend.position="bottom",
        text = element_text(size=10))

plotbools

#From the plot, we donot want the field with more than 330 NA values
boolNA <- sample_health_bool_counts %>% filter(!(answers %in% c("TRUE","FALSE"))) %>% filter(n > 300) %>% select(question) %>% pull
```
```{r}
#cate and AB
fields_cate <- fieldsinfo %>% filter(data_type=="categorical" | Scale=="AB") %>% select(field) %>% pull

sample_health_cate <- sample_health %>% select(Family_ID, any_of(fields_cate))

# get the counts for true, false and NA answers for each questions
sample_health_cate_long <- sample_health_cate %>% pivot_longer(!Family_ID, names_to="question", values_to="answers")
sample_health_cate_counts <- sample_health_cate_long %>% group_by(question, answers) %>% tally()

# plot the distribution
plotcate <- sample_health_cate_counts %>% ggplot(aes(x=question, y=n, fill=answers)) +
  geom_bar(stat="identity")+
  coord_flip() +
  theme(legend.position="bottom",
        text = element_text(size=10))

plotcate

#From the plot, we don't want the field with more than 300 NA values
cateNA <- sample_health_cate_counts %>% filter(is.na(answers)) %>% filter(n > 300) %>% select(question) %>% pull
```

###Creat SAMPLE and complete physeq
```{r}
sample_health <- sample_health %>% select(!any_of(boolNA)) %>% select(!any_of(cateNA))
SAMPLE = sample_data(as.data.frame(sample_health))
physeq = phyloseq(OTU, TAX, SAMPLE)
```




#Plot

##PCoA
###overview all infant sample with all sample_type
```{r}
#Subset infant samples
#Define detection and prevalence thresholds to filter out rare taxa
pseq.infant <- subset_samples(physeq, !(sample_type %in% c("F","M"))) %>% core(detection = 0.1/100, prevalence = 1/100)

set.seed(4235421)
ord <- ordinate(pseq.infant, "MDS", "bray")
```

```{r}
#basic descriptions
plot_ordination(pseq.infant, ord, color = "sample_type") +
  geom_point(size = 5)

plot_ordination(pseq.infant, ord, color = "inf_DeliveryMode") +
  geom_point(size = 5)

plot_ordination(pseq.infant, ord, color = "Run_ID") +
  geom_point(size = 5)
```


bool variables
```{r}
# Illustrate sample similarities with PCoA (NMDS)
#%>% filter(!(field %in% c("inf_HasDiagPollenAllergies")))
#%>% filter(!(field %in% boolNA))
fields_bool <- fieldsinfo %>% filter(data_type=="bool") %>% filter(!(field %in% boolNA)) %>% select(field) %>% pull

for (item in fields_bool) {
  p <- plot_ordination(pseq.infant, ord, color = item) +
    geom_point(size = 5) +
    labs(title = item)
  pbool <- try(print(p), silent = T)
  if (grepl("Error", pbool)) {print(item)}
  else {print(p)}
}
``` 

categorical and AB scaled variables
```{r}
fields_cate <- fieldsinfo %>% filter(data_type=="categorical" | Scale=="AB") %>% select(field) %>% pull

for (item in fields_cate) {
  p <- plot_ordination(pseq.infant, ord, color = item) +
    geom_point(size = 5)
  print(p)
}
```
##Plot Function
for PCoA, there might be many subphyseq and variables.
Here is a function to do this
```{r}
phy.pcoa <- function(dataset, fields) {
  set.seed(4235421)
  ord <- ordinate(dataset, "MDS", "bray")
  for (item in fields) {
    p <- plot_ordination(dataset, ord, color = item) +
      geom_point(size = 5)
    pbool <- try(print(p), silent = T)
    if (grepl("Error", pbool)) {print(item)}
    else {print(p)}
    }
}
```
```{r}
#B1
pseq.pcoa <- subset_samples(physeq, (sample_type %in% c("B5"))) %>% core(detection = 0.1/100, prevalence = 1/100)

phy.pcoa(pseq.pcoa, c("inf_HadVaccine"))
```

#PERMANOVA
```{r}
#here generate the subseq
# %>% subset_samples(!is.na(colorname.str))

colorname.str <- "inf_RegDefecation"

pseq.infant <- subset_samples(physeq, !sample_type %in% c("F","M")) %>% subset_samples(!is.na(get(colorname.str)))

pseq.rel <- pseq.infant


#dram landscape plot
pseq.rel <- microbiome::transform(pseq.rel, "compositional")
otu <- abundances(pseq.rel)
meta <- meta(pseq.rel)

p <- plot_landscape(pseq.rel, method = "NMDS", distance = "bray", col = colorname.str, size = 5)
print(p)
```
```{r}
#p-value
permanova <- adonis(t(otu) ~ inf_RegDefecation,
               data = meta, permutations=99, method = "bray")
print(as.data.frame(permanova$aov.tab)[1, "Pr(>F)"])

#homogeneity retest
dist <- vegdist(t(otu))

anova(betadisper(dist, meta$inf_RegDefecation))[1,"Pr(>F)"]
```
```{r}
#Top taxa
bott <- gsub(" ", "",paste(colorname.str,"1"))
coef <- coefficients(permanova)[bott,]
top.coef <- coef[rev(order(abs(coef)))[1:20]]
par(mar = c(3, 14, 2, 1))
title <- paste(colorname.str,"Top taxa")
barplot(sort(top.coef), horiz = T, las = 1, main = title)
```

Here is a for loop to calculate p-values of all variables.
```{r}

fields_sample <- fieldsinfo %>% filter(data_type %in% c("bool","categorical") | Scale=="AB") %>% filter(!(field %in% boolNA)) %>% filter(!(field %in% cateNA)) %>% select(field) %>% pull

pvalue.frame <- data.frame(variables=c(), p_value=c(), homogeneity_pvalue=c())

for (item in fields_sample) {
  pseq.pvalue <- subset_samples(physeq, !sample_type %in% c("F","M")) %>% subset_samples(!is.na(get(item))) %>% core(detection = 0.1/100, prevalence = 1/100)
  pseq.rel <- microbiome::transform(pseq.pvalue, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- try(adonis(t(otu) ~ get(item),data = meta, permutations=99, method = "bray"),silent = T)
  if ('try-error' %in% class(permanova)) {
    print(c(item, "Error"))
    next
  } else {
    #pvalue
    pvalue <- as.data.frame(permanova$aov.tab)[1, "Pr(>F)"]
    dist <- vegdist(t(otu))
    
    #homogeneity retest
    hpvalue <- anova(betadisper(dist, meta[[item]]))[1,"Pr(>F)"]
    newrow <- data.frame(variables=item, p_value=pvalue, homogeneity_pvalue=hpvalue)
    pvalue.frame <- rbind(pvalue.frame,newrow)
    
    #plot
    p <- plot_landscape(pseq.rel, method = "NMDS", distance = "bray", col = item, size = 5) +
      labs(title = paste(item,pvalue))
    print(p)
  }
}

```

here is a for loop to plot top taxa
```{r}

fields_pvalue <- pvalue.frame %>% filter(p_value <0.05) %>% select(variables) %>% pull
for (item in fields_sample) {
  pseq.pvalue <- subset_samples(physeq, !sample_type %in% c("F","M")) %>% subset_samples(!is.na(get(item))) %>% core(detection = 0.1/100, prevalence = 1/100)
  pseq.rel <- microbiome::transform(pseq.pvalue, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- try(adonis(t(otu) ~ get(item),data = meta, permutations=99, method = "bray"),silent = T)
  if ('try-error' %in% class(permanova)) {
    print(c(item, "Error"))
    next
  } else {
    #taxa
    coef <- coefficients(permanova)[2,]
    top.coef <- coef[rev(order(abs(coef)))[1:20]]
    par(mar = c(3, 14, 2, 1))
    pvalue <- as.data.frame(permanova$aov.tab)[1, "Pr(>F)"]
    title <- paste(item,"pvalue=",pvalue,"--Top taxa")
    barplot(sort(top.coef), horiz = T, las = 1, main = title)
  }
}
```

