---
title: "Metaphlan_Health_3"
output: html_notebook
---
This scrip merged health data with metaphlan3 data

```{r warning=FALSE, include=FALSE}
# create Phylloseq object
library(ape)
library(dplyr)
library(phyloseq)
library(tidyverse)
library(microbiome)
library(vegan)

# import tables and format for PhylloSeq
metaphlan <- read_tsv("../Metaphlan3/merged_abundance_table.txt") 
# select the species level information
metaphlan_s <- metaphlan %>% separate(clade_name,  #separate clade_name into 7 columns
                                    into=c("kingdom","Phylum","class","order","family","genera","species"), 
                                    sep="\\|") %>%
                            separate(NCBI_tax_id, #separete NCBI_tax_id into 7 columns
                                    into=c("Tax_kingdom","Tax_Phylum","Tax_class","Tax_order","Tax_family","Tax_genera","TaxID"), 
                                    sep="\\|") %>%
                            mutate(kingdom=str_replace(kingdom, "k__", "")) %>%
                            mutate(Phylum=ifelse(kingdom=="Unknown", "Unknown", str_replace(Phylum, "p__", ""))) %>%
                            mutate(class=ifelse(kingdom=="Unknown", "Unknown", str_replace(class, "c__", ""))) %>%
                            mutate(order=ifelse(kingdom=="Unknown", "Unknown", str_replace(order, "o__", ""))) %>%
                            mutate(family=ifelse(kingdom=="Unknown", "Unknown", str_replace(family, "f__", ""))) %>%
                            mutate(genera=ifelse(kingdom=="Unknown", "Unknown", str_replace(genera, "g__", ""))) %>%
                            mutate(species=ifelse(kingdom=="Unknown", "Unknown", str_replace(species, "s__", ""))) %>%
                            mutate(TaxID=ifelse(kingdom=="Unknown", "-1", TaxID)) %>%
                            #columns from clade_name: keep only the tax-name and trim the "k__" prefix
                            filter(!is.na(species))
                            #filter out the rows that have no species information.

# remove samples with less than 10 taxonomic assignments
nonzero <- function(x) sum(x != 0)
meta_exclude <- metaphlan_s %>% summarise(across(starts_with("Hel"), nonzero)) %>% 
  pivot_longer(everything(),names_to="samples", values_to="nb_tax") %>%
  filter(nb_tax<10) %>% select(samples) %>%
  pull()

metaphlan_s <- metaphlan_s %>% select(- all_of(meta_exclude))

## create TaxTable
taxMat <-  metaphlan_s %>% select(TaxID, kingdom, Phylum, class, order, family, genera, species) 
taxMat <- taxMat %>% tibble::column_to_rownames("TaxID")
TAX = tax_table(as.matrix(taxMat))

## create countTable
countMat <- metaphlan_s %>% select(-kingdom, -Phylum, -class, -order, -family, -genera, -species, 
                                   -Tax_kingdom, -Tax_Phylum, -Tax_class, -Tax_order, -Tax_family, -Tax_genera)
countMat <- countMat %>% tibble::column_to_rownames("TaxID")
OTU = otu_table(as.matrix(countMat), taxa_are_rows = TRUE)
```

Now we have TAX OTU but no SAMPLE.

```{r include=FALSE}
## sample description file
sampleMat <- read_csv("../Metaphlan3/sample_table_12.04.21.csv")

health_3m_meta <- read.csv("../Health_data/Health_3months_nonOutliers_ABscaled_17.04.21.csv")
health_Reg_meta <- read.csv("../Health_data/Health_Reg_nonOutliers_ABscaled_17.04.21.csv")

fieldsinfo <- read.csv("../Health_data/fields_usable_21.04.15.csv")

#Here I want to add all fields except "text" type fields from health_data into SAMPLE
#For late analysis, use info from 'fieldsinfo' to cluster the fields groups

fields_text <- fieldsinfo %>% filter(data_type=="text") %>% select(field) %>% pull
health_3m <- health_3m_meta %>% select(!any_of(fields_text))
health_Reg <- health_Reg_meta %>% select(!any_of(fields_text))

sample_health_3m <- left_join(sampleMat, health_3m, by = c("Family_ID", "sample_type"))
sample_health_Reg <- left_join(sampleMat, health_Reg, by = c("Family_ID", "sample_type"))
```

#Merge 3M data
There are two extra rows for Hel178xB8 and Hel126xB9.
The number of obs. of sample_health_3m_Mat should be the same as sampleMat(374)
```{r include=FALSE}
#Valid_from for B8 and B9
#B8: 52(keep) 65
#B9: 78(keep) 91
sample_health_3m_Mat <- sample_health_3m %>% filter(!(Valid_From %in% c(65,91)))
```

#Merge Reg data
To delete the rows with the same familyID and time code,
we need to create a list contain all suspect rows by Valid_from/to
and a list without any suspect rows called cleaned list.

Some not wanted rows might remain in cleaned list. And some are misfiltered for having the same Valid_from/to.

Then with anti_join we are able to gain a new list having misfiltered rows.

Finally, we can merge the misfiltered rows with the cleaned list by full_join.

The number of obs. of sample_health_Reg_Mat should be the same as sampleMat(374)

Note: in Reg_data, there are more than 2 B1 for one family.
And we need to treat mis_sample as the samole_health_Reg with the valide_from of secondary priority.
Thus the code here is not perfect but works for this job.

```{r include=FALSE}
#Valid_from for each time code
#Priority B1 B2 B3 B4 B5 B6 B7 B8 B9 F M
#1st      2 5 8 11 22 30 43 48 NA NA NA
#2nd      1 4 7 10 20 26 39 NA NA NA NA
#3rd      0 3 6 9 16 24 35 NA NA NA NA

clean_sample <- sample_health_Reg %>% filter(Valid_From %in% c(2,5,8,11,22,30,43,48))
suspect_sample <- sample_health_Reg %>% filter(!(Valid_From %in% c(2,5,8,11,22,30,43,48)))
mis_sample <- anti_join(suspect_sample,clean_sample, by = c("Family_ID", "sample_type"))

#filter mis_sample with 2nd priority valid_from
mis_clean <- mis_sample %>% filter(Valid_From %in% c(1, 4, 7, 10, 20, 26, 39))
mis_sus <- mis_sample %>% filter(!Valid_From %in% c(1, 4, 7, 10, 20, 26, 39))
mis_mis <- anti_join(mis_sus, mis_clean, by = c("Family_ID", "sample_type") )
mis_filtered <- full_join(mis_clean, mis_mis)

sample_health_Reg_Mat <- full_join(clean_sample, mis_filtered)
#the number of obs. of sample_health_Reg_Mat is 376, which means there are two extra rows.
#The extra rows can be found manually: "Hel655xB5" "Hel655xB5"
#Family 655 have three B5 valid from 18,13,12
#here use valid_from == 18
sample_health_Reg_Mat <- sample_health_Reg_Mat %>% filter(!(Family_ID=="Perhe655" & Valid_From %in% c(13,12)))

```

#SAMPLE
Now merge 3m_Mat and Reg_Mat to creat SAMPLE
```{r}
#Merge both 3m_Mat and Reg_Mat
sample_health <- left_join(sample_health_3m_Mat, sample_health_Reg_Mat, by = c("Family_ID", "sample_type","sample_ID","file_name","inf_DeliveryMode", "Run_ID"))

#creat SAMPLE
sample_health <- sample_health %>% tibble::column_to_rownames("sample_ID")
SAMPLE = sample_data(as.data.frame(sample_health))

physeq = phyloseq(OTU, TAX, SAMPLE)
```

#Plot

```{r}
#Subset infant samples
#Define detection and prevalence thresholds to filter out rare taxa
pseq.infant <- subset_samples(physeq, !(sample_type %in% c("M","F"))) %>% core(detection = 0.1/100, prevalence = 1/100)

set.seed(4235421)
ord <- ordinate(pseq.infant, "MDS", "bray")
```

bool variables
```{r}
# Illustrate sample similarities with PCoA (NMDS)
fields_bool <- fieldsinfo %>% filter(data_type=="bool") %>% filter(!(field %in% c("inf_HasDiagPollenAllergies"))) %>% select(field) %>% pull

for (item in fields_bool) {
  p <- plot_ordination(pseq.infant, ord, color = item) +
    geom_point(size = 5)
  print(p)
}
```

categorical and AB scaled variables
```{r}
fields_cate <- fieldsinfo %>% filter(data_type=="categorical" | Scale=="AB") %>% select(field) %>% pull

for (item in fields_cate) {
  p <- plot_ordination(pseq.infant, ord, color = item) +
    geom_point(size = 5)
  print(p)
}
```



```{r}

pseq.general <- subset_samples(physeq, !is.na(inf_InfectiousDisease)) %>%
   aggregate_taxa(level = "Phylum")

p <- plot_composition(pseq.general, group_by = "inf_InfectiousDisease")
print(p)
```

