---
title: "Metaphlan_Health_3"
output: html_notebook
---
This scrip merged health data with metaphlan3 data

```{r echo=FALSE, warning=FALSE}
# create Phylloseq object
library(ape)
library(dplyr)
library(phyloseq)
library(tidyverse)
library(microbiome)
library(vegan)

# import tables and format for PhylloSeq
metaphlan <- read_tsv("../Metaphlan3/merged_abundance_table.txt") 
# select the species level information
metaphlan_s <- metaphlan %>% separate(clade_name,  #separate clade_name into 7 columns
                                    into=c("kingdom","Phylum","class","order","family","genera","species"), 
                                    sep="\\|") %>%
                            separate(NCBI_tax_id, #separete NCBI_tax_id into 7 columns
                                    into=c("Tax_kingdom","Tax_Phylum","Tax_class","Tax_order","Tax_family","Tax_genera","TaxID"), 
                                    sep="\\|") %>%
                            mutate(kingdom=str_replace(kingdom, "k__", "")) %>%
                            mutate(Phylum=ifelse(kingdom=="Unknown", "Unknown", str_replace(Phylum, "p__", ""))) %>%
                            mutate(class=ifelse(kingdom=="Unknown", "Unknown", str_replace(class, "c__", ""))) %>%
                            mutate(order=ifelse(kingdom=="Unknown", "Unknown", str_replace(order, "o__", ""))) %>%
                            mutate(family=ifelse(kingdom=="Unknown", "Unknown", str_replace(family, "f__", ""))) %>%
                            mutate(genera=ifelse(kingdom=="Unknown", "Unknown", str_replace(genera, "g__", ""))) %>%
                            mutate(species=ifelse(kingdom=="Unknown", "Unknown", str_replace(species, "s__", ""))) %>%
                            mutate(TaxID=ifelse(kingdom=="Unknown", "-1", TaxID)) %>%
                            #columns from clade_name: keep only the tax-name and trim the "k__" prefix
                            filter(!is.na(species))
                            #filter out the rows that have no species information.

# remove samples with less than 10 taxonomic assignments
nonzero <- function(x) sum(x != 0)
meta_exclude <- metaphlan_s %>% summarise(across(starts_with("Hel"), nonzero)) %>% 
  pivot_longer(everything(),names_to="samples", values_to="nb_tax") %>%
  filter(nb_tax<10) %>% select(samples) %>%
  pull()

metaphlan_s <- metaphlan_s %>% select(- all_of(meta_exclude))

## create TaxTable
taxMat <-  metaphlan_s %>% select(TaxID, kingdom, Phylum, class, order, family, genera, species) 
taxMat <- taxMat %>% tibble::column_to_rownames("TaxID")
TAX = tax_table(as.matrix(taxMat))

## create countTable
countMat <- metaphlan_s %>% select(-kingdom, -Phylum, -class, -order, -family, -genera, -species, 
                                   -Tax_kingdom, -Tax_Phylum, -Tax_class, -Tax_order, -Tax_family, -Tax_genera)
countMat <- countMat %>% tibble::column_to_rownames("TaxID")
OTU = otu_table(as.matrix(countMat), taxa_are_rows = TRUE)
```

Now we have TAX OTU but no SAMPLE.

```{r echo=FALSE}
## sample description file
sampleMat <- read_csv("../Metaphlan3/sample_table_12.04.21.csv")
health3m_meta <- read.csv("../Health_data/Health_3months_nonOutliers_ABscaled_17.04.21.csv")
fieldsinfo <- read.csv("../Health_data/fields_usable_21.04.15.csv")

#add only certain fields from 3M into sample description file
fieldsAB <- fieldsinfo %>% filter(Scale=="AB") %>% select(field) %>% pull
health3m <- health3m_meta %>% select(Family_ID,sample_type,Valid_From,Valid_to,any_of(fieldsAB))

sample_health_Mat <- left_join(sampleMat, health3m, by = c("Family_ID", "sample_type"))
```
To delete the rows with the same familyID and time code,
we need to create a list contain all suspect rows by Valid_from/to
and a list without any suspect rows called cleaned list.
Some of these rows are not wanted but some are misfiltered for having the same Valid_from/to
Then with anti_join we are able to gain a new list having misfiltered rows
Finally, we can merge the misfiltered rows with the cleaned list by full_join

(This step is a little complex because I thought some family have two B3 valid from 39 and 52 respectly, but some family may have only one B3 valid from 39 or 52. If drop all B3 valid from 39 then families have only one B3 will be cleaned by mistake. But the truth is there is no such a family...)
```{r}
#Valid_from for each time code
#B1: -1 or 0
#B2 has no repeat rows: 26
#B3: 52
#B4:has no repeat rows: 65
#B5: 91
#no valid health data for B6 B7 B8 B9 F M
clean_sample <- sample_health_Mat %>% filter(Valid_From %in% c(-1,0,26,52,65,91))
suspect_sample <- sample_health_Mat %>% filter(!(Valid_From %in% c(-1,0,26,52,65,91)))
mis_sample <- anti_join(suspect_sample,clean_sample, by = c("Family_ID", "sample_type"))
sample_health_3 <- full_join(clean_sample, mis_sample)

#creat SAMPLE
sample_health3 <- sample_health_3 %>%tibble::column_to_rownames("sample_ID")
SAMPLE = sample_data(as.data.frame(sample_health3))

physeq = phyloseq(OTU, TAX, SAMPLE)
```
```{r}
# Define detection and prevalence thresholds to filter out rare taxa
physeq.core <- core(physeq, detection = 0.1/100, prevalence = 1/100)

set.seed(4235421)
ord <- ordinate(physeq.core, "MDS", "bray")

# Illustrate sample similarities with PCoA (NMDS)
plot_ordination(physeq.core, ord, color = "familly_WorriedHealth") +
  geom_point(size = 5)

plot_ordination(physeq.core, ord, color = "inf_DiaperRashIntensity") +
  geom_point(size = 5)

plot_ordination(physeq.core, ord, color = "inf_FoldingRashIntensity") +
  geom_point(size = 5)

plot_ordination(physeq.core, ord, color = "inf_GeneralHealth") +
  geom_point(size = 5)

plot_ordination(physeq.core, ord, color = "inf_HealthSameAge") +
  geom_point(size = 5)

plot_ordination(physeq.core, ord, color = "inf_InfectiousDisease") +
  geom_point(size = 5)

plot_ordination(physeq.core, ord, color = "inf_SeriousIllness") +
  geom_point(size = 5)

plot_ordination(physeq.core, ord, color = "inf_SkinStatus") +
  geom_point(size = 5)
```
```{r}
pseq.general <- subset_samples(physeq, !is.na(inf_InfectiousDisease)) %>%
   aggregate_taxa(level = "Phylum")

p <- plot_composition(pseq.general, group_by = "inf_InfectiousDisease")
print(p)
```

